### 2021-07-22 paper/paper_draft.tex
data redo complete

\newcommand{\insitu}{\textit{in situ} }

@@

We supplement these data with a variety of other information.  Most importantly, we add facility-level compliance data from Alberta's greenhouse gas emissions pricing policies, the SGER (2009-2017) and the CCIR (2018-2019). These data were provided for this analysis by the Government of Alberta, with some portions of the data also available publicly. We supplement these data with information from the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for smaller facilities.\footnote{The federal data are useful since the reporting threshold of 10kt \coe per annum is lower than the provincial threshold for coverage under the emissions pricing policies of 100kt\coe per annum.}  Since these data are both reported at the facility level, not by generating unit, we apply these measurements in the same way in our paper and so we cannot differentiate between multiple generating units on particular sites.\footnote{There are two exceptions to this. The Keephills 3 and Genessee 3 coal-fired generating units report separately under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.}  Since we have only annual data, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units.

@@

Finally, we supplement our publicly-available data with several series from NRGStream, a commercial data aggregation service. Most importantly, this service provides daily natural gas prices for the Nova Inventory Transfer hub which are a better proxy for Alberta natural gas prices than publicly-available series for Henry Hub. NRGStream also scrapes the real-time data from the AESO, and thus provides slightly different data from the aggregate, publicly-available historic metered volumes provided by the grid operator. We use NRGStream real-time data to compile real-time intertie flows between Alberta and British Columbia, Saskatchewan, and Montana and to source real-time generation (as opposed to net-to-grid metered volumes) from all generating units in the province where needed.

All data save the natural gas prices and NRGStream generation and trade date which we are not authorized to redistribute are available HERE and can be replicated using code for the paper which is available at GITHUB LINK. For the publicly available code, we substitute Henry Hub gas prices as a placeholder, to which we apply the average discount observed between Henry Hub and NIT to adjust the values so that regression results may be more closely replicated.

@@

As Figure \ref{fig:sector_compliance_cost} shows, changes in facility compliance costs were not all coincident with increased marginal carbon prices, but are actually dominated by the changes in output-based allocations which occurred in 2018-2019 rather than the initial increase carbon prices from \$15 to \$30 per tonne \coe between 2016 and 2018. The largest cost increases affect coal-fired power plants and some simple-cycle generators, while have limited effects on combined heat and power plants. Compliance costs for natural gas combined-cycle plants dropped between 2017 and 2019 as a result of a change in the formula by which output-based allocations were calculated.

@@

The next stage in our analysis is to transfer our information on facility compliance costs to the synthetic merit orders. Recall that, for the major emitters, we have specific facility-level information on their emissions-intensity their output-based allocations of emissions credits which determines their net emissions pricing liabilities. Since we are interested in the separate effects of both components of emissions policies, and since both vary at the facility level through our sample, we can separately identify the impacts of each on price pass-through behaviour. As such, for each of the synthetic merit orders we create for analysis purposes, we also capture and store both the marginal emissions charge and the marginal output-based allocation at each percentile support point.

@@

To these, synthetic merit orders, we add our hourly-level data on market status, weather, and fuel prices specified above. This then allows us to build regression equations for each percentile level of each synthetic merit order, asking the degree to which carbon price and output-based allocation changes impact the level of each step of the supply curves by facility, plant-type, by owner, or for the complete market.

>>>>

\newcommand{\insitu}{\textit{in situ} }
\newcommand{\SGER}{\textit{SGER} }
\newcommand{\CCIR}{\textit{CCIR} }

@@

We supplement these data with a variety of other information.  Most importantly, we add facility-level compliance data from Alberta's greenhouse gas emissions pricing policies, the \SGER (2009-2017) and the \CCIR (2018-2019). These data were provided for this analysis by the Government of Alberta, with some portions of the data also available publicly. We supplement these data with information from the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for smaller facilities.\footnote{The federal data are useful since the reporting threshold of 10kt \coe per annum is lower than the provincial threshold for coverage under the emissions pricing policies of 100kt\coe per annum.}

There are three sources of incompleteness in our emissions data. First, we have no emissions data for smaller facilities, since provincial and federal reporting is not required below a minimum threshold. We have data on all facilities that fell under Alberta's industrial carbon pricing programs, but for smaller facilities which would have been subject only to the economy-wide carbon pricing program imposed in 2018, we have limited information. Federal reporting thresholds decreased to 10,000 tonnes per year which translates to an exemption from reporting for most natural gas generators with installed capacity of less than 20MW. Because reporting thresholds have changed over time, in some cases we have partial information on emissions intensities for smaller facilities, while in other cases we have no information at all. We use emissions-intensities from provincial reporting data as our first option, since the provincial reporting includes both verified production and emissions. Where we don't have provincial data, we use generation data aggregated by facility and federally-reported emissions to calculate an emissions intensity. Where we have partial information, we complete our dataset by filling first backward in time and then forward in time by plant by year such that, for example, a plant for which we only have 2019 data would see their 2019 emissions intensity used for every previous year. Where we have no information at all, we assume an emissions-intensity of 0.55 tonnes per MWh for simple cycle natural gas plants, the only type of plant for which we lack information of this type, based on facilities for which we do have information in our sample. The second source of incompleteness comes from the fact that data are reported at the facility level, not by generating unit. We apply the same emissions intensity to multiple generating units within a facility.\footnote{There are two exceptions to this. The Keephills 3 and Genessee 3 coal-fired generating units report separately under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.} Since we have only annual data at the facility level, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units. Finally, combined heat and power units present a challenge since many units report their emissions for the entire production facility, not simply for the power plant. There are some stand-alone units, but these also vary in terms of the emissions-intensity of net-to-grid electricity depending on the design of the specific unit and what share of the produced heat is used to generate electricity vs process steam. The emissions reporting data tell us the total electricity and heat produced in these facilities, and the emissions attributed to them, but do not tell us what the emissions intensity of an incremental unit of electricity would be. The output-based allocations for cogeneration facilities are determined by the carbon pricing policies, so we know the value of these for each regulated entity. For the purposes of our analysis, we adopt an assumed emissions intensity of electricity of 0.418 tonnes per MWh for combined heat and power units. Biomass plants, as well as wind, solar and hydroelectric plants are assigned a deemed emissions intensity of zero.

The compliance data also identifies two other attributes of importance to our analysis. First, under the \SGER, facilities only faced compliance costs after their third full year of operations. Because facilities had to report federally during these periods, we can identify emissions intensities for some large facilities for 3 or 4 years before they face carbon pricing, and then for the remaining years in the sample when they are subject to carbon pricing. This is confounded slightly by the fact that the output-based allocations in future years are a function of the emissions intensity in the first years of operation. We account for this with an indicator variable in our analysis for the pre-compliance period. The data also provide the facility-specific rate at which output-based allocations were provided in each year, and the carbon price faced on emissions net of output-based allocations. The carbon price is constant for all Alberta-based facilities, but the combined-cycle plant in Fort Nelson, BC faces the BC carbon tax while being connected to the Alberta electricity grid, and so receives no output-based allocations and faced higher carbon prices in many years of the sample than comparable plants in Alberta.

@@

Finally, we supplement our publicly-available data with several series from NRGStream, a commercial data aggregation service. Most importantly, this service provides daily natural gas prices for the Nova Inventory Transfer hub which are a better proxy for Alberta natural gas prices than publicly-available series for Henry Hub. NRGStream also scrapes the real-time data from the AESO, and thus provides slightly different data from the aggregate, publicly-available historic metered volumes provided by the grid operator. We use NRGStream real-time data to compile real-time intertie flows between Alberta and British Columbia, Saskatchewan, and Montana and to source real-time generation (as opposed to net-to-grid metered volumes) from all generating units in the province where needed. These data are generally the same as metered volumes with the exception of cases where an industrial production facility is co-located with the power generation facility, in which case only the net-to-grid volumes from the facility are measured in the metered volumes data while the NRGStream scraped data captures actual production by hour for the generating facility.

All data save the natural gas prices and NRGStream generation and trade data which we are not authorized to redistribute are available HERE and can be replicated using code for the paper which is available at GITHUB LINK. For the publicly available code, we substitute Henry Hub gas prices as a placeholder, to which we apply the average discount observed between Henry Hub and NIT to adjust the values so that regression results may be more closely replicated.

@@

As Figure \ref{fig:sector_compliance_cost} shows, changes in facility compliance costs were not all coincident with increased marginal carbon prices, but are actually dominated by the changes in output-based allocations which occurred subsequent to the adoption of the CCIR in 2018 rather than the initial increase carbon prices from \$15 to \$30 per tonne \coe in 2016 and 2017. The largest cost increases affect coal-fired power plants and some simple-cycle generators, while limited effects were felt by combined heat and power plants. Compliance costs for natural gas combined-cycle plants dropped between 2017 and 2019 as a result of a change in the formula by which output-based allocations were calculated.

@@

The next stage in our analysis is to transfer our information on facility compliance costs to the synthetic merit orders. Recall that, for the major emitters, we have facility-year-specific information on their emissions-intensity and their output-based allocations of emissions credits which determines their net emissions pricing liabilities. Since we are interested in the separate effects of both components of emissions policies, and since both vary at the facility level through our sample, we can separately identify the impacts of each on price pass-through behaviour. As such, for each of the synthetic merit orders we create for analysis purposes, we also capture and store both the marginal emissions charge and the marginal output-based allocation at each percentile support point.

@@

To these, synthetic merit orders, we add our hourly data on market status, weather, and fuel prices specified above. This then allows us to build regression equations for each percentile support point of each synthetic merit order, asking the degree to which carbon price and output-based allocation changes impact the level of each step of the supply curves by facility, plant-type, by owner, or for the complete market.
